---
title: "2022 - day 02"
author: "Hendrik Wagenseil"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false

library(reticulate)
use_virtualenv(
  file.path(
    here::here()
    , ".venv"
  )
)

knitr::opts_knit$set(
  root.dir = '~/Documents/repos/github/adventofcode'
)
```

## Data 

Load and preview the data set. 

```{r}
#| label: data-load

y = "2022"
d = "02"

ifl = file.path(
  here::here()
  , y
  , "data"
  , paste0(y, "-day", d, "input")
)

txt = readLines(ifl)
head(txt, n = 30L)
```

## Part I

Get the score of the rock paper scissors game.

* A = rock, B = paper, C = scissors
* X = rock, Y = paper, Z = scissors

::: {.panel-tabset}

#### R 

Approach:  

* define a named vector `pts` with all possible combinations and the 
resulting score
* match `txt` with `pts` to get the score for all games and sum up

```{r}
pts = c(
  "A X" = 1 + 3
  , "A Y" = 2 + 6
  , "A Z" = 3 + 0
  , "B X" = 1 + 0
  , "B Y" = 2 + 3
  , "B Z" = 3 + 6
  , "C X" = 1 + 6
  , "C Y" = 2 + 0
  , "C Z" = 3 + 3
)

score = pts[match(txt, table = names(pts))]
sum(score)
```

#### Python 

Approach:  


```{python}
txt = r.txt

pts = {
  'A X': 1 + 3
  , 'A Y': 2 + 6
  , 'A Z': 3 + 0
  , 'B X': 1 + 0
  , 'B Y': 2 + 3
  , 'B Z': 3 + 6
  , 'C X': 1 + 6
  , 'C Y': 2 + 0
  , 'C Z': 3 + 3
}

score = [pts[i] for i in txt]
sum(score)
```

:::


## Part II

Get the score of the rock paper scissors game.

* A = rock, B = paper, C = scissors
* X = loose, Y = draw, Z = win

::: {.panel-tabset}

#### R 

Approach:  

* same as in part I, just re-define vector `pts` 

```{r}
pts = c(
  "A X" = 3 + 0
  , "A Y" = 1 + 3
  , "A Z" = 2 + 6
  , "B X" = 1 + 0
  , "B Y" = 2 + 3
  , "B Z" = 3 + 6
  , "C X" = 2 + 0
  , "C Y" = 3 + 3
  , "C Z" = 1 + 6
)

score = pts[match(txt, table = names(pts))]
sum(score)
```

#### Python 

Approach:  

* same as in part I, just re-define vector `pts` 

```{python}
txt = r.txt

pts = {
  'A X': 3 + 0
  , 'A Y': 1 + 3
  , 'A Z': 2 + 6
  , 'B X': 1 + 0
  , 'B Y': 2 + 3
  , 'B Z': 3 + 6
  , 'C X': 2 + 0
  , 'C Y': 3 + 3
  , 'C Z': 1 + 6
}

score = [pts[i] for i in txt]
sum(score)
```

:::
